---
title: "wk7_prac"
output: html_document
date: "2025-11-25"
---

Library and downloading data
```{r}
library(here)
library(janitor)
library(sf)
library(tidyverse)
library(tmap)
library(spdep)
library(RColorBrewer)
#read the ward data in
LondonWards <- st_read("/Users/levine/Desktop/CASA/GIS/wk1/statistical-gis-boundaries-london/ESRI/London_Ward.shp")

```


```{r}
LondonWardsMerged <- st_read("/Users/levine/Desktop/CASA/GIS/wk1/statistical-gis-boundaries-london/ESRI/London_Ward_CityMerged.shp")%>%
  st_transform(.,27700)

```
```{r}
WardData <- read_csv("https://data.london.gov.uk/download/f33fb38c-cb37-48e3-8298-84c0d3cc5a6c/772d2d64-e8c6-46cb-86f9-e52b4c7851bc/ward-profiles-excel-version.csv",
                    locale = locale(encoding = "latin1"),
                     na = c("NA", "n/a")) %>% 
  clean_names()

LondonWardsMerged <- LondonWardsMerged %>% 
  left_join(WardData, 
            by = c("GSS_CODE" = "new_code"))%>%
  dplyr::distinct(GSS_CODE, .keep_all = T)%>%
  dplyr::select(GSS_CODE, ward_name, average_gcse_capped_point_scores_2014)
```
Check if it's in the right projection
```{r}
#have a look to check that it's 
#in the right projection
st_crs(LondonWardsMerged)
```


Plot the data to check is it reasonable
```{r}
library(tmap)

BluePlaques <- st_read("https://s3.eu-west-2.amazonaws.com/openplaques/open-plaques-london-2018-04-08.geojson") %>%
    st_transform(.,27700)
```
```{r}
tmap_mode("plot")
tm_shape(LondonWardsMerged) +
  tm_polygons(fill_alpha = 0.5) +
tm_shape(BluePlaques) +
  tm_dots(fill = "blue", size=0.1)
```

Lose the plaques that outside of London
```{r}
summary(BluePlaques)
```
```{r}
BluePlaquesSub <- BluePlaques[LondonWardsMerged,]

tmap_mode("plot")
```
```{r}
tm_shape(LondonWardsMerged) +
  tm_polygons(fill_alpha = 0.5) +
tm_shape(BluePlaquesSub) +
  tm_dots(fill = "blue", size=0.1)
```


Count all of the blue plaques that fall within each Ward in the City
```{r}
example<-st_intersects(LondonWardsMerged, BluePlaquesSub)

example
```

We can see that the polygon with the ID of 6 Kingston upon Thames - Coombe Hill has three plaques within it. Check it with st_join.
```{r}
check_example <- LondonWardsMerged%>%
  st_join(BluePlaquesSub)%>%
  filter(ward_name=="Kingston upon Thames - Coombe Hill")
```


Calculate the length of each list in each polygon and add it as new column.
```{r}
library(sf)
points_sf_joined <- LondonWardsMerged%>%
  mutate(n = lengths(st_intersects(., BluePlaquesSub)))%>%
  janitor::clean_names()%>%
  #calculate area
  mutate(area=st_area(.))%>%
  #then density of the points per ward
  mutate(density=n/area)%>%
  #select density and some other variables 
  dplyr::select(density, ward_name, gss_code, n, average_gcse_capped_point_scores_2014)
```

a quick choropleth map
```{r}
points_sf_joined<- points_sf_joined %>%                    
  group_by(gss_code) %>%         
  summarise(density = first(density),
          wardname= first(ward_name),
          plaquecount= first(n))

tm_shape(points_sf_joined) +
    tm_polygons(
      fill = "density",
      fill.scale = tm_scale_intervals(
        values = "brewer.blues",
        style="jenks"),
      # set the legend
      fill.legend = tm_legend(title="Blue Plaque Density",
                              title.size=0.85,
                              size=0.8,
                              # plot outside of the main map
                              #explained below
                              position=tm_pos_out("right", 
                                                  "center",
                                                  pos.v = "center")))
```

define a Wij spatial weights matrix before calculate Moran’s I and any similar statistics.
```{r}
library(spdep)
```

First calculate the centroids of all Wards in London
```{r}
#First calculate the centroids of all Wards in London

coordsW <- points_sf_joined%>%
  st_centroid()%>%
  st_geometry()
  
plot(coordsW,axes=TRUE)
```

Generate a spatial weights matrix.

First create a neibourhood list use queen case:
```{r}
#create a neighbours list
LWard_nb <- points_sf_joined %>%
  poly2nb(., queen=T)

```

```{r}
summary(LWard_nb)
```
Average neibours is 5.888.

Now plot the neibours:
```{r}
#plot them
plot(LWard_nb, st_geometry(coordsW), col="red")
#add a map underneath
plot(points_sf_joined$geometry, add=T)
```


Now make a spatial weight matrix from the weight list.
```{r}
#create a spatial weights matrix from these weights
Lward.lw <- LWard_nb %>%
  nb2mat(., style="B")

sum(Lward.lw)
```

Standardization:where 1 is divided by the sum of the number of neighbours in each row
```{r}
sum(Lward.lw[1,])
```

```{r}
Lward.lw <- LWard_nb %>%
  nb2listw(., style="C")
```

Moran’s I test tells us whether we have clustered values (close to 1) or dispersed values (close to -1). Calculate the density rather than the raw value.  neighbours are set by the weight matrix.
1 = clustered, 0 = no pattern, -1 = dispersed
```{r}
I_LWard_Global_Density <- points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  moran.test(., Lward.lw)

I_LWard_Global_Density
```

Geary's C can find whether similar values or dissimilar values are clustering.
1 means no spatial autocorrelation, <1 - positive spatial autocorrelation or similar values clustering, >1 - negative spatial autocorreation or dissimilar values clustering
```{r}
C_LWard_Global_Density <- 
  points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  geary.test(., Lward.lw)

C_LWard_Global_Density
```

Getis Ord:
```{r}
G_LWard_Global_Density <- 
  points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  globalG.test(., Lward.lw)

G_LWard_Global_Density
```

Local version of Moran'I
```{r}
#use the localmoran function to generate I for each ward in the city

I_LWard_Local_count <- points_sf_joined %>%
  pull(plaquecount) %>%
  as.vector()%>%
  localmoran(., Lward.lw)%>%
  as_tibble()

I_LWard_Local_Density <- points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  localmoran(., Lward.lw)%>%
  as_tibble()

#what does the output (the localMoran object) look like?
slice_head(I_LWard_Local_Density, n=5)
```

copy some of the columns (the I score (column 1) and the z-score standard deviation (column 4)) back into the LondonWards spatialPolygonsDataframe
```{r}
points_sf_joined <- points_sf_joined %>%
  mutate(plaque_count_I = as.numeric(I_LWard_Local_count$Ii))%>%
  mutate(plaque_count_Iz =as.numeric(I_LWard_Local_count$Z.Ii))%>%
  mutate(density_I =as.numeric(I_LWard_Local_Density$Ii))%>%
  mutate(density_Iz =as.numeric(I_LWard_Local_Density$Z.Ii))
```


plot a map of the local Moran’s I outputs
```{r}
breaks1<-c(-1000,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)
```

create a new diverging colour brewer palette and reverse the order
```{r}
library(RColorBrewer)
MoranColours<- rev(brewer.pal(7, "RdGy"))
```

Plot the output:
```{r}
tm_shape(points_sf_joined) +
    tm_polygons(
      fill = "plaque_count_Iz",
      fill.scale = tm_scale_intervals(values = "brewer.br_bg",
                                      midpoint = "NA",
                                      style = "fixed",
                                      breaks=breaks1),
       fill.legend = tm_legend(title="Local Moran's I, Blue Plaques in London",
                              title.size=0.85,
                              size=0.8,
                              # plot outside of the main map
                              #explained below
                              position=tm_pos_out("right", 
                                                  "center",
                                                  pos.v = "center")))
```

```{r}
Gi_LWard_Local_Density <- points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  localG(., Lward.lw)

head(Gi_LWard_Local_Density)
```

Add the Getis Ord G∗i data to the simple featur
```{r}
points_sf_joined <- points_sf_joined %>%
  mutate(density_G = as.numeric(Gi_LWard_Local_Density))
```

And the map of output
```{r}
tm_shape(points_sf_joined) +
    tm_polygons(
      fill = "density_G",
      fill.scale = tm_scale_intervals(values = "brewer.br_bg",
                                      midpoint = "NA",
                                      style = "fixed",
                                      breaks=breaks1),
       fill.legend = tm_legend(title="Gi*, Blue Plaques in London",
                              title.size=0.85,
                              size=0.8,
                              # plot outside of the main map
                              #explained below
                              position=tm_pos_out("right", 
                                                  "center",
                                                  pos.v = "center")))
```

Have a look at other variables:
GCSE scores
```{r}
#use head to see what other variables are in the data file

slice_head(points_sf_joined, n=2)
```

print the class of each column (But need to drop the grometry)
```{r}
Datatypelist <- LondonWardsMerged %>% 
  st_drop_geometry()%>%
  summarise_all(class) %>%
  pivot_longer(everything(), 
               names_to="All_variables", 
               values_to="Variable_class")

Datatypelist
```

```{r}
I_LWard_Local_GCSE <- LondonWardsMerged %>%
  arrange(GSS_CODE)%>%
  pull(average_gcse_capped_point_scores_2014) %>%
  as.vector()%>%
  localmoran(., Lward.lw)%>%
  as_tibble()

points_sf_joined <- points_sf_joined %>%
  arrange(gss_code)%>%
  mutate(GCSE_LocIz = as.numeric(I_LWard_Local_GCSE$Z.Ii))

tm_shape(points_sf_joined) +
    tm_polygons(
      fill = "GCSE_LocIz",
      fill.scale = tm_scale_intervals(values = "brewer.br_bg",
                                      midpoint = "NA",
                                      style = "fixed",
                                      breaks=breaks1),
       fill.legend = tm_legend(title="Local Moran's I, GCSE Scores",
                              title.size=0.85,
                              size=0.8,
                              # plot outside of the main map
                              #explained below
                              position=tm_pos_out("right", 
                                                  "center",
                                                  pos.v = "center")))
```

Now the Gi* statistic to look at clusters of high and low scores and explain what the output map is showing
```{r}
G_LWard_Local_GCSE <- LondonWardsMerged %>%
  dplyr::arrange(GSS_CODE)%>%
  dplyr::pull(average_gcse_capped_point_scores_2014) %>%
  as.vector()%>%
  localG(., Lward.lw)

points_sf_joined <- points_sf_joined %>%
  dplyr::arrange(gss_code)%>%
  dplyr::mutate(GCSE_LocGiz = as.numeric(G_LWard_Local_GCSE))


tm_shape(points_sf_joined) +
    tm_polygons(
      fill = "GCSE_LocGiz",
      fill.scale = tm_scale_intervals(values = "brewer.br_bg",
                                      midpoint = "NA",
                                      style = "fixed",
                                      breaks=breaks1),
       fill.legend = tm_legend(title="Gi*, GCSE Scores",
                              title.size=0.85,
                              size=0.8,
                              # plot outside of the main map
                              #explained below
                              position=tm_pos_out("right", 
                                                  "center",
                                                  pos.v = "center")))
```

